---
title: "Tidy Data Mini-Lab"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
    css: ../../../styles.css
  pdf_document: default
---

**Duration: 30 minutes** 

Now it's your turn to practice the verbs on a new dataset. Complete the following task:

# MVP

**Task 1.**  
Load in the `country_data.csv` file and take a look at it. What needs changed in this to make it tidy? Which columns need made into long format, and which need spread to rows?

```{r}
country <- read_csv("data/country_data.csv")
```
<br>

**Task 2.**  
Do the appropriate pivoting of your columns using R  

```{r}

# we need to make the year longer and the type wider.

country_longer <- pivot_longer(country,
  cols = Y1999:Y2001,
  names_to = "year",
  values_to = "measuremrnts"
)

country_wider <- country_longer %>% 
pivot_wider(
            names_from = "type",
            values_from = "measuremrnts")
   
country_longer  
country_wider

```



<br>

**Task 3.**  
Separate the `country` variable so you have the name in one column and the number in another. 

```{r}
country_separed <- country_wider %>% 
  separate(country, c("number_id", "name"), sep = "\\.")

country_separed

```


<br>

**Task 4.**  
Unite the `code_letter` and `code_numb` variables so that they are both in one column, with no spaces separating them.   

```{r}

country_unite <- country_separed %>% 
  unite(code_id, c("code_letter", "code_numb"), sep = ":")

country_unite
```
5. reordering the columns

```{r}
country_ordered <- select(country_unite, code_id, name, number_id, year, population, cases)

country_ordered
```

