---
title: "weekend_homework_week_1"
output: html_document
date: "2023-02-25"
---
```{r}
library(tidyverse)
library(dplyr)
library(lubridate)
library(ggplot2)
```
1. investigating the data: what are the variable names, variable types, dimensions, how many missing values, the columns name and the head of our dataser.
```{r}
book_csv <- read_csv('data/books.csv')
glimpse(book_csv)
dim(book_csv)
names(book_csv)
sum(is.na(book_csv))
head(book_csv)

```
2. create a new table without (rowid, isbn13, bookID, isbn) 


```{r}
 books <- book_csv %>% 
  select(-rowid, -isbn13, -bookID, -isbn)  

```

3. what are the book with highest average_rating? 
```{r}
books %>% 
slice_max(average_rating)
# this reveals that 22 books are rated 5.
```

4. what are the books with average_rating higher than the total_average_rating
```{r}
books %>% 
mutate(total_average_rating = mean(average_rating), .after = average_rating) %>%
  mutate(above_total_average_rating = (average_rating > total_average_rating),
         .after = total_average_rating ) %>% 
    filter(above_total_average_rating == TRUE)  
# this code reveals that 5,960 books are rated above the total avg rating.

```


5. plot the 5 authors with highest number of books published.

```{r}
 best_publishing_authors <- books %>% 
group_by(authors) %>% 
summarise(nber_of_books_puplished = n()) %>% 
slice_max(nber_of_books_puplished, n = 5)
ggplot(best_publishing_authors, aes(x = authors,y = nber_of_books_puplished, fill = authors)) + geom_col(stat='identity') +
labs( title = "Best Publishing Authors")


best_publishing_authors


```

number of Publishers used by the Stephen King
```{r}
books %>% 
  select(title, authors, publisher) %>% 
  filter(authors == "Stephen King") %>% 
group_by(publisher)
# Stephen King has written 40 books and they were published by 26 publishers.
```

6. plot the top 10 most rated books and their authors.

```{r}
most_rated_books <- books %>% 
    select(title, authors, num_pages,average_rating, ratings_count) %>% 
slice_max(ratings_count, n = 10) 

ggplot(most_rated_books, aes(x = title, y = ratings_count, , fill = authors)) + geom_col(stat='identity') +theme(axis.text.x = element_text(angle = 90)) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 35))
                                                                                                                                        
```
4. nber of book published by each publisher. top 10

```{r}
books %>% 
  group_by(publisher) %>% 
summarise(number_book = n()) %>% 
slice_max(number_book, n = 10)
 # the above code gives the top 10 publisher in this data set.

books %>% 
  group_by(authors) 
# this code reveals the number of authors in this data set, which is 6,639.

books %>% 
group_by(publisher)
   # the above code gives the total number of publishers in this data set.


  
```

5. authors of the top publisher, Vintage
```{r}
books %>% 
  select(title, authors, publisher) %>% 
  filter(publisher == "Vintage") %>% 
group_by(authors)

# from this code we see that Vintage has published books 318 books for 203 different authors.

```
6. ploting the books by year
```{r}
books_by_year<- books %>% 
  mutate(year = format(as.Date(books$publication_date, format="%m/%d/%Y"),"%Y"), .after = publication_date) %>% 
group_by(year) %>%   # here we can see that these books were published between 1900 and 2020
  group_by(authors) 
  #summarise(Number_of_book_per_year = n())

  ggplot(books_by_year, aes(x = year, y = authors)) + geom_point() 
labs( title = "books published between 1900 and 2020")

####from jamie

books_by_year <- books %>%
  mutate(year = format(as.Date(books$publication_date,
                               format="%m/%d/%Y"), "%Y"),
         .after = publication_date) 

  # group_by(year) %>%   # here we can see that these books were published between 1900 and 2020
  # # group_by(authors)
  # summarise(Number_of_book_per_year = n()) %>%
  # ungroup()

  ggplot(books_by_year, aes(x = year)) + 
    geom_histogram(stat = "count") +
    labs(title = "books published between 1900 and 2020") 
    theme(axis.text.x = element_text(angle = 90))
    

    
    ######from david
    
books_by_year<- books %>%
  mutate(
    year = year(as.Date(publication_date, format="%m/%d/%Y")), .after = publication_date)

ggplot(books_by_year) +
  aes(x = year) +
  geom_histogram(colour = "white") +
  labs(
    x = "",
    y = "Number of Books",
    title = "Number of Books per year",
    subtitle = "in GoodReads Data"


```

7. what is the year with most book published?

```{r}
books %>% 
  mutate(year = format(as.Date(books$publication_date, format="%m/%d/%Y"),"%Y"), .after = publication_date) %>% 
group_by(year) %>%   # here we can see that these books were published between 1900 and 2020
  summarise(Number_of_book_per_year = n()) %>% 
slice_max(Number_of_book_per_year)
```
8. sorting

```{r}
books %>% 
  arrange(desc(average_rating), desc(title))


books %>% 
  arrange(desc(average_rating), title)
```
```{r}
 best_publishing_authors <- books %>% 
group_by(authors) %>% 
summarise(nber_of_books_puplished = n()) %>% 
slice_max(nber_of_books_puplished, n = 5)
ggplot(best_publishing_authors, 
       aes(x = nber_of_books_puplished,y = average_rating)) + geom_point() +
 facet_wrap(~ authors)

```

