---
title: "Extra `dplyr` Practice"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
    css: ../../../styles.css
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<br>
The dataset for this minilab is from [Kaggle](https://www.kaggle.com/zynicide/wine-reviews) and contains reviews of different wine varieties (location, winery, price).

# MVP

```{r}
library(readr)
library(dplyr)
library(tidyr)

```


<br>

**Question 1.** 

Read in the data and investigate it, e.g. variables, variable types, dimensions etc.

```{r}
wine_rating <- read_csv('data/wine_ratings.csv')
dim(wine_rating)
glimpse(wine_rating)
head(wine_rating, 10)
```


<br>

**Question 2.** 

Select columns `title`, `taster_name`, `points`, `price`, and `country`, then **drop all NAs**. Save into a variable called `wine_subset`. Use this variable for all subsequent tasks.

```{r}
wine_subset <- drop_na(select(wine_rating, title, taster_name, points, price, country))
```


<br>

**Question 3.** 

Find the 5 countries with the highest mean points.

```{r}
wine_subset %>% 
  group_by(country) %>% 
  summarise(mean_points = mean(points)) %>% 
  #arrange(desc(mean_points))
slice_max(mean_points, n = 5)
 

```


<br>

**Question 4.** 

Find the 5 countries with the lowest mean prices.

```{r}
wine_subset %>% 
  group_by(country) %>% 
  summarise(mean_price = mean(price)) %>% 
  # arrange(mean_price)
  slice_min(mean_price, n = 5)
```


<br>

**Question 5.** 

Does the result differ if you use the median instead of the mean for Questions 3 and 4?

```{r}  
 # Q3 using median  shows a different result, the top 5 countries are not the same
wine_subset %>% 
  group_by(country) %>% 
  summarise(median_point = median(points)) %>% 
  arrange(desc(median_point))

# Q4 using median also show a different result, the top 5 countries are not the same

wine_subset %>% 
  group_by(country) %>% 
  summarise(median_price = median(price)) %>% 
  arrange(median_price)

 
```


<br>

**Question 6.** 

Find the names of all the tasters using `distinct`. Then take a random sample of 5 names using `sample_n`. Save these 5 names into a new variable called `taster_sample` which should be a **character vector** containing only the names. 

<details>
<summary>**Hint**</summary>: You can use the function `pull` to extract a single column as a vector.
</details>

```{r}
#wine_subset %>% 
  #distinct( taster_name, keep_all = TRUE) %>% 
  #sample_n(5) %>% 
  #pull(var = 1, name = taster_name)

taster_sample <- wine_subset %>% 
  distinct( taster_name) %>% 
  sample_n(5) %>% 
  pull()

taster_sample
```


<br>

**Question 7**

Do the same thing for `country`, i.e. create a **character vector** called `country_sample` using `distinct` and `pull`, but this time use `sample_frac` to pick 10 countries at random. 

<details>
<summary>**Hint**</summary>: What is the proportion of countries you want? $\frac{Part}{Whole}$
</details>

```{r}
country_sample <- wine_subset %>% 
  distinct(country) %>% 
  sample_n(10) %>% 
  #sample_frac(10/n()) %>% 
  pull()
  
country_sample
  
```


<br>

**Question 8**

Filter the rows of `wine_subset` so you are left only with data from the 5 tasters in your `taster_sample` and from the countries in your `country_sample`. Save this into a new variable called `wine_sample`.

```{r}
  wine_sample <- filter(  wine_subset, taster_name %in% taster_sample)
wine_sample

 country_sample <- filter(  wine_subset, country %in% country_sample)
country_sample


```


<br>

**Question 9**

Using your `wine_sample`, create a new column `difference_from_avg` which contains the difference between each wine's rating (`points`) and the average rating for that taster, then arrange the differences to see the highest positive difference first. [**Hint:** you will first have to calculate the average points for each taster]

```{r}
wine_sample_new <- wine_sample %>% 
  group_by(taster_name) %>% 
  mutate(taster_avg_points = mean(points)) %>% 
  mutate(difference_from_avg = points - taster_avg_points,
         .before = points) %>% 
  arrange(desc(difference_from_avg))
wine_sample_new
  
```



