---
title: "Meteorites Weekend Homework"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
    css: ../../../styles.css
  pdf_document: default
---
```{r}
library(tidyverse)
```


# MVP

The goal of this homework is to produce a reproducible analysis of some meteorites. Rather than work solely in a Notebook, we're going to write an R cleaning script as well. By the end you'll create:

1. An .R file that cleans some data on meteorites.
2. A .Rmd file that takes your cleaned data and performs some analysis.

You will also write a short `README.md` that describes the structure of your project.

The data you want to clean is in the file called `meteorite_landings.csv`. It comes from NASA and has information on meteorites that have been found up to the year 2013. 

## The .R file

To clean the data please do the following.

1. Read the data into R
```{r}
meteorite <- read_csv("data/meteorite_landings.csv") 
meteorite <- janitor::clean_names(meteorite)
```

2. Change the names of the variables to follow our naming standards.
```{r}
meteorite <- janitor::clean_names(meteorite)
```

3. Split the column `GeoLocation` into latitude and longitude, the new latitude and longitude columns should be numeric.
```{r}
meteorite_split <- meteorite %>% 
  mutate( geo_location = gsub("[()]", "", meteorite$geo_location)) %>% 
  separate(col = "geo_location", into = c("latitude" , "longitude"), sep = ", ")  
  
  
meteorite_split
  
```


```{r}
meteorite_split <- meteorite_split %>% 
mutate( latitude = as.numeric(latitude)) %>% 
  mutate(longitude = as.numeric(longitude))

```


4. Replace any missing values in latitude and longitude with zeros.

```{r}
meteorite_replaced <- meteorite_split %>% 
  #filter(is.na(longitude))
  mutate(latitude = coalesce(latitude, 0)) %>% 
   mutate(longitude = coalesce(longitude, 0)) 

```

5. Remove meteorites less than 1000g in weight from the data.
```{r}
meteorite_not_less_than_1000 <- meteorite_replaced %>% 
  filter(mass_g >= 1000)
  
  
```

6. Order the data by the year of discovery.
```{r}
meteorite_ordered <- meteorite_not_less_than_1000 %>% 
  arrange(desc(year))
```


## The .Rmd file

In the .Rmd file please do the following.

1. Read the cleaned data into your environment.

```{r}
meteorite_replaced

```


2. Find the names and years found for the 10 largest meteorites in the data.
```{r}

meteorite_largest <- meteorite_replaced %>% 
  arrange(desc(mass_g)) %>% 
 head(10) %>% 
  select(name, year)
  


```


3. Find the average mass of meteorites that were recorded falling, vs. those which were just found.
```{r}
meteorite_replaced_avg <- meteorite_replaced %>% 
  #filter(is.na(mass_g))
mutate(mass_g = coalesce(mass_g, median(mass_g, na.rm = TRUE)))  
  # filter(is.na(mass_g))

meteorite_replaced_avg %>% 
  mutate(average_fell = mean(subset(meteorite_replaced_avg, fall== "Fell")$ mass_g)) %>% 
 mutate(average_found = mean(subset(meteorite_replaced_avg, fall== "Found")$ mass_g)) %>% 
  summarise ( average_fell ,  average_found) 

  
```

4. Find the number of meteorites in each year, for every year since 2000.

```{r}
meteorite_by_year <-  meteorite_replaced %>% 
  filter(year >= 2000) %>% 
  group_by(year) %>% 
  summarise(year_count = n())  


 
  year1m <- meteorite_by_year$year_count [meteorite_by_year$year == 2000]
  
  class(meteorite_by_year)

```

5. Write a function called `get_meteorite_count()` that takes in two arguments: a dataframe and a year, and returns the phrase "In (input_year), there were (number of meteorites that year) meteorites." IF there are 1 or more meteorites that year. If there are no meteorites in the year, return the phrase "As far as we know, there were no meteorites that year".

```{r}
 get_meteorite_count <- function(dataframe, Y) {
   
   nber_meteorite <-  dataframe$year_count [dataframe$year == Y]
   
  
     return( paste(c("In ",Y," there were ", nber_meteorite, " meteorites."), collapse = " "))
  
   
   return( "As far as we know, there were no meteorites that year")}
  
 

get_meteorite_count(meteorite_by_year,2006)


```



## The README.md file

Make a file called `README.md`. In this file we just want you to describe the structure of your project, and write a general summary of your code.

